---
output:
  title: "Other geographical units in the prescribing dataset"
  author: "Data Science in Biomedicine - Assessment"
  date: "`r Sys.Date()`"
  pdf_document: default
---
# Setup Chunks

These chunks are as follows:
- R Markdown setup chunk
- Load libraries required for execution
- Create two vectors - one with a list of URLs of all datasets since October 2015 from the NHS Scotland Open data website. The other contains a list of values representing the month and year of the datasets in the format mmyyyy. This full chunk is not included as it would take up most of the report!

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r LoadLibraries, message=FALSE}
library(tidyverse)
library(lubridate)
library(readxl)
library(ggplot2)
library(gt)
library(purrr)
library(here)
```

# Data Import and Wrangle

## Setup custom functions

### Camel
This uses the `clean_names()` function from `janitor` but implements it with CamelCase rather than snake_case.

```{r}
Camel <- function(Dataset) {
  Dataset %>% 
  clean_names(., "upper_camel")
}
```

### Wrangle

This function will perform an analysis pipeline on the data in a DRY way.

```{r WrangleFunction}
Wrangle <- function(DatasetToWrangle)
{
  DatasetToWrangle <- DatasetToWrangle %>%
    filter(GPPractice %in% GPPractices$GPPractice) %>%
    mutate(HB = case_when(is.na(HBT) ~ HBT2014, !is.na(HBT) ~ HBT)) %>%
    select(-c(
      HBT2014,
      HBT,
      ClassOfPreparationCode,
      NumberOfPaidItems,
      GrossIngredientCost
    ))
  DatasetToWrangle <-
    full_join(DatasetToWrangle, HealthBoards, by = "HB")
  DatasetToWrangle <-
    full_join(DatasetToWrangle, GPPractices, by = "GPPractice")
  DatasetToWrangle <- DatasetToWrangle %>%
    mutate(HB = HB.x)
  DatasetToWrangle <-
    full_join(DatasetToWrangle, SIMD, by = c("DataZone", "HB"))
  DatasetToWrangle <- DatasetToWrangle %>%
    select(-c(HB.x, HB.y, DZ, HBcode)) %>%
    drop_na() %>%
    mutate(HealthBoard = HB) %>%
    mutate(SIMDOverallRank = SIMD2020v2_Vigintile) %>%
    mutate(SIMDHealthRank = SIMD2020_Health_Domain_Rank) %>%
    mutate(UrbanRuralCode = URclass) %>%
    mutate(UrbanRuralDescription = URname) %>%
    select(-c(
      HB,
      SIMD2020v2_Vigintile,
      SIMD2020_Health_Domain_Rank,
      URclass,
      URname
    )) %>%
    mutate(HealthBoard = as.factor(HealthBoard)) %>%
    mutate(SIMDOverallRank = as.factor(SIMDOverallRank)) %>%
    mutate(SIMDHealthRank = as.factor(SIMDHealthRank)) %>%
    mutate(UrbanRuralCode = as.factor(UrbanRuralCode)) %>%
    mutate(UrbanRuralDescription = as.factor(UrbanRuralDescription)) %>%
    mutate(Postcode = as.factor(Postcode)) %>%
    mutate(DataZone = as.factor(DataZone)) %>%
    mutate(HBName = as.factor(HBName)) %>%
    mutate(HealthBoardCode = as.factor(HealthBoardCode)) %>%
    mutate(GPPractice = as.factor(GPPractice)) %>% 
    mutate(PaidDateMonth = lubridate::ymd(PaidDateMonth)) %>% 
    mutate(Month = lubridate::month(PaidDateMonth)) %>%
    mutate(Year = lubridate::year(PaidDateMonth))
  
  return(DatasetToWrangle)
}
```

## Prescriptions in the Community Data

This chunk builds a tibble of the two vectors above. It also includes some use of `lubridate` functions to perform some data wrangling.

```{r URLWrangle}
URLList <- read_csv("https://raw.githubusercontent.com/s1906007/expert-octo-pancake/main/Data/URLList.csv") %>% 
  mutate(YearMonths = ym(YearMonths)) %>%
  mutate(Month = month(YearMonths)) %>%
  mutate(Year = year(YearMonths))
```

This chunk imports all the data from the NHS Scotland Open Data website and stores each month as an R dataframe. This will take an hour or more to run!

```{r PullAllDataFiles}
URLList <- URLList %>% 
  mutate(MonthYear = paste0(Year, Month))

for (row in 1:nrow(URLList)) {
  assign(paste0("DF", URLList$MonthYear[row]), read_csv(URLList$URLs[row]))
  row = row + 1
}
```

```{r RegExBindDataFrames}
DFList <- mget(ls(pattern = "DF20"))
CompletePrescriptionDataset <- bind_rows(DFList)
rm(list = ls(pattern = "DF"))
```

```{r ReadOrWriteRDS, eval=FALSE, include=FALSE}
#writeRDS(CompletePrescriptionDataset, here("Data", "CompletePrescriptionDataset.RDS"))
CompletePrescriptionDataset <- readRDS(here("Data", "CompletePrescriptionDataset.RDS"))
```

## Other Datasets

We'll also import three more datasets. Health Board names and codes, GP Practice data and codes, and the Scottish Index of Multiple Deprivation data from 2020.

```{r LoadAndWrangleOtherDatasets}
#Health Boards
HBCodes <- read_csv("https://raw.githubusercontent.com/s1906007/expert-octo-pancake/main/Data/HBCodes.csv")

HealthBoards <-
  read_csv(
    "https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/652ff726-e676-4a20-abda-435b98dd7bdc/download/hb14_hb19.csv"
  )
HealthBoards <- HealthBoards %>%
  select(c(HB, HBName))
HealthBoards <-
  full_join(HealthBoards, HBCodes %>% select(-HBName), by = c("HB"))

#Scottish Index of Multiple Deprivation
download.file(
  "https://www.gov.scot/binaries/content/documents/govscot/publications/statistics/2020/01/scottish-index-of-multiple-deprivation-2020-data-zone-look-up-file/documents/scottish-index-of-multiple-deprivation-data-zone-look-up/scottish-index-of-multiple-deprivation-data-zone-look-up/govscot%3Adocument/SIMD%2B2020v2%2B-%2Bdatazone%2Blookup.xlsx",
  (here("Data", "SIMD.xlsx"))
)

SIMD <- read_excel((here("Data", "SIMD.xlsx")), sheet = 3)

SIMD <- SIMD %>%
  select(
    c(
      DZ,
      SIMD2020v2_Vigintile,
      SIMD2020_Health_Domain_Rank,
      Population,
      URclass,
      URname,
      HBcode
    )
  ) %>%
  mutate(DataZone = DZ) %>%
  mutate(HB = HBcode)

#GP Practices
GPPractices <-
  read_csv(
    "https://www.opendata.nhs.scot/dataset/f23655c3-6e23-4103-a511-a80d998adb90/resource/1a15cb34-fcf9-4d3f-ad63-1ba3e675fbe2/download/practice_contactdetails_oct2022-open-data.csv"
  )

GPPractices <- GPPractices %>%
  select(c(PracticeCode, PracticeListSize, Postcode, HB, DataZone)) %>%
  mutate(GPPractice = PracticeCode) %>% 
  select(-PracticeCode)

#Add COMMUNITY PHARMACY row
CommunityPharmacies <- list(5295403, "ALL SCO", "S92000003", "S92000003", 99998)

GPPractices <- rbind(GPPractices, CommunityPharmacies)


#Numbers of women who would be elegible for contraceptive prescriptions
download.file(
  "https://www.scotlandscensus.gov.uk/media/rfzlmc3l/scotland-blk.zip",
  here("Data", "scotland-blk.zip")
)
AgeBySex <-
  read_csv(unzip(here("Data", "scotland-blk.zip"), files = "DC1117SC.csv", exdir = here("Data")))
ReproActive <- AgeBySex %>%
  mutate(Age = ...2) %>%
  select(Age, Females) %>%
  filter(between(Age, 16, 49))

ReproNum <- sum(ReproActive[["Females"]])
```

# Contraception

## Short-acting

### Tablets

```{r TabletsData}
Tablet <- CompletePrescriptionDataset %>%
  filter(str_detect(BNFItemCode, "0703010")) %>%
  filter(str_detect(BNFItemDescription, "TAB"))

Tablet <- bind_rows(Tablet, CompletePrescriptionDataset %>% filter(str_detect(BNFItemCode, "0703021")))
```

## LARC

### Injections

```{r InjectionsData}
Injection <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemCode, "0703022M"))
```

### IUS

```{r IUSData}
IUS <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemCode, "0703023"))
```

### IUD

```{r IUDData}
IUD <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemCode, "21040"))
```

### Implants

```{r ImplantsData}
Implant <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemDescription, "NEXPLANON"))
```

## Emergency Contraception

```{r EmergencyContraceptionData}
EmergencyContraception <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemCode, "0703050"))
```


```{r WrangleContraceptivesData}
EmergencyContraception <- Wrangle(EmergencyContraception)
Implant <- Wrangle(Implant)
Injection <- Wrangle(Injection)
IUD <- Wrangle(IUD)
IUS <- Wrangle(IUS)
Tablet <- Wrangle(Tablet)
```

```{r LARCJoins}
LARC <- full_join(Implant, Injection)
LARC <- full_join(LARC, IUD)
LARC <- full_join(LARC, IUS)
```





Good to here!


```{r WranglingAndPlottingLARC}
LARCSmall <- LARC %>% 
  select(HB, PaidQuantity, PaidDateMonth) %>% 
  mutate(HB = as.factor(HB)) %>% 
  mutate(PaidDateMonth = as.factor(PaidDateMonth)) %>%
  group_by(HB, PaidDateMonth) %>% 
  summarise(sum(PaidQuantity)) %>% 
  pivot_wider(names_from = HB, values_from = `sum(PaidQuantity)`)
  

LARCSmall %>% 
  ggplot(aes(x = PaidDateMonth, y = S08000015)) +
  geom_point() 

  facet_wrap(facets = vars(HB))
```

```{r}
library(scales)

ggplot(df, aes(x=days, y=value, color=factor(year(date)))) +
  geom_line()+scale_x_date(labels = date_format("%b"))

LARCGroup <- LARC %>%
  group_by(across(c(HealthBoardCode, PaidDateMonth)))
summarise(across(sum(PaidQuantity))) %>%
  ggplot(
    aes(x = PaidDateMonth, y = sum(PaidQuantity)) +
      geom_point() +
      facet_wrap( ~ year(PaidDateMonth), scales = "free") +
      scale_x_date(date_labels = "%b"
      ))

  
# Works
LARC %>% 
  filter(HealthBoardCode == "AaA") %>% 
  ggplot(aes(x = PaidDateMonth, y = PaidQuantity)) +
  geom_point() +
  scale_x_date()
```

