---
output:
  pdf_document:
      includes:
          in_header: header.tex 
---
# Setup Chunks

These chunks are as follows:
- R Markdown setup chunk
- Load libraries required for execution
- Create two vectors - one with a list of URLs of all datasets since October 2015 from the NHS Scotland Open data website. The other contains a list of values representing the month and year of the datasets in the format mmyyyy. This full chunk is not included as it would take up most of the report!

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r LoadLibraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(readxl)
library(ggplot2)
library(gt)
library(purrr)
library(here)
library(colorspace)
library(patchwork)
```

# Data Import and Wrangle

## Setup custom function - Wrangle

This function will perform an analysis pipeline on the data in a DRY way.

```{r WrangleFunction}
Wrangle <- function(DatasetToWrangle)
{
  DatasetToWrangle <- DatasetToWrangle %>%
    filter(GPPractice %in% GPPractices$GPPractice) %>%
    mutate(HB = case_when(is.na(HBT) ~ HBT2014, !is.na(HBT) ~ HBT)) %>%
    select(-c(
      HBT2014,
      HBT,
      ClassOfPreparationCode,
      NumberOfPaidItems,
      GrossIngredientCost
    ))
  DatasetToWrangle <-
    full_join(DatasetToWrangle, HealthBoards, by = "HB")
  DatasetToWrangle <-
    full_join(DatasetToWrangle, GPPractices, by = "GPPractice")
  DatasetToWrangle <- DatasetToWrangle %>%
    mutate(HB = HB.x)
  DatasetToWrangle <-
    full_join(DatasetToWrangle, SIMD, by = c("DataZone", "HB"))
  DatasetToWrangle <- DatasetToWrangle %>%
    select(-c(HB.x, HB.y, DZ, HBcode)) %>%
    drop_na() %>%
    mutate(HealthBoard = HB) %>%
    mutate(SIMDOverallRank = SIMD2020v2_Vigintile) %>%
    mutate(SIMDHealthRank = SIMD2020_Health_Domain_Rank) %>%
    mutate(UrbanRuralCode = URclass) %>%
    mutate(UrbanRuralDescription = URname) %>%
    select(-c(
      HB,
      SIMD2020v2_Vigintile,
      SIMD2020_Health_Domain_Rank,
      URclass,
      URname
    )) %>%
    mutate(HealthBoard = as.factor(HealthBoard)) %>%
    mutate(SIMDOverallRank = as.factor(SIMDOverallRank)) %>%
    mutate(SIMDHealthRank = as.factor(SIMDHealthRank)) %>%
    mutate(UrbanRuralCode = as.factor(UrbanRuralCode)) %>%
    mutate(UrbanRuralDescription = as.factor(UrbanRuralDescription)) %>%
    mutate(Postcode = as.factor(Postcode)) %>%
    mutate(DataZone = as.factor(DataZone)) %>%
    mutate(HBName = as.factor(HBName)) %>%
    mutate(HealthBoardCode = as.factor(HealthBoardCode)) %>%
    mutate(GPPractice = as.factor(GPPractice)) %>% 
    mutate(PaidDateMonth = ymd(PaidDateMonth))
  
  return(DatasetToWrangle)
}
```

```{r}
CompletePrescriptionDataset <- readRDS(here
                                       ("Data", 
                                         "CompletePrescriptionDataset.RDS"))
```

## Other Datasets

We'll also import three more datasets. Health Board names and codes, GP Practice data and codes, and the Scottish Index of Multiple Deprivation data from 2020.

```{r LoadAndWrangleOtherDatasets, message=FALSE, warning=FALSE}
#Health Boards
HBCodes <- read_csv("https://raw.githubusercontent.com/s1906007/expert-octo-pancake/main/Data/HBCodes.csv")

HealthBoards <-
  read_csv(
    "https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/652ff726-e676-4a20-abda-435b98dd7bdc/download/hb14_hb19.csv"
  )
HealthBoards <- HealthBoards %>%
  select(c(HB, HBName))
HealthBoards <-
  full_join(HealthBoards, HBCodes %>% select(-HBName), by = c("HB"))

#Scottish Index of Multiple Deprivation
download.file(
  "https://www.gov.scot/binaries/content/documents/govscot/publications/statistics/2020/01/scottish-index-of-multiple-deprivation-2020-data-zone-look-up-file/documents/scottish-index-of-multiple-deprivation-data-zone-look-up/scottish-index-of-multiple-deprivation-data-zone-look-up/govscot%3Adocument/SIMD%2B2020v2%2B-%2Bdatazone%2Blookup.xlsx",
  (here("Data", "SIMD.xlsx"))
)

SIMD <- read_excel((here("Data", "SIMD.xlsx")), sheet = 3)

SIMD <- SIMD %>%
  select(
    c(
      DZ,
      SIMD2020v2_Vigintile,
      SIMD2020_Health_Domain_Rank,
      Population,
      URclass,
      URname,
      HBcode
    )
  ) %>%
  mutate(DataZone = DZ) %>%
  mutate(HB = HBcode)

#GP Practices
GPPractices <-
  read_csv(
    "https://www.opendata.nhs.scot/dataset/f23655c3-6e23-4103-a511-a80d998adb90/resource/1a15cb34-fcf9-4d3f-ad63-1ba3e675fbe2/download/practice_contactdetails_oct2022-open-data.csv"
  )

GPPractices <- GPPractices %>%
  select(c(PracticeCode, PracticeListSize, Postcode, HB, DataZone)) %>%
  mutate(GPPractice = PracticeCode) %>% 
  select(-PracticeCode)

#Add COMMUNITY PHARMACY row
CommunityPharmacies <- list(5295403, "ALL SCO", "S92000003", "S92000003", 99998)

GPPractices <- rbind(GPPractices, CommunityPharmacies)

#Numbers of women who would be elegible for contraceptive prescriptions
download.file(
  "https://www.scotlandscensus.gov.uk/media/rfzlmc3l/scotland-blk.zip",
  here("Data", "scotland-blk.zip")
)
AgeBySex <-
  read_csv(unzip(here("Data", "scotland-blk.zip"), files = "DC1117SC.csv", exdir = here("Data")))
ReproActive <- AgeBySex %>%
  mutate(Age = ...2) %>%
  select(Age, Females) %>%
  filter(between(Age, 16, 49))

ReproNum <- sum(ReproActive[["Females"]])
```

# Contraception Data

```{r ContraceptionData, message=FALSE, warning=FALSE}
#Tablets
Tablet <- CompletePrescriptionDataset %>%
  filter(str_detect(BNFItemCode, "0703010")) %>%
  filter(str_detect(BNFItemDescription, "TAB"))

Tablet <- bind_rows(Tablet, CompletePrescriptionDataset %>% filter(str_detect(BNFItemCode, "0703021")))

#Injections
Injection <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemCode, "0703022M"))

#IUS
IUS <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemCode, "0703023"))

#IUD
IUD <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemCode, "21040"))

#Implants
Implant <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemDescription, "NEXPLANON"))

#Emergency Contraception
EmergencyContraception <- CompletePrescriptionDataset %>% 
  filter(str_detect(BNFItemCode, "0703050"))

#Wrangle Contraceptive Data
EmergencyContraception <- Wrangle(EmergencyContraception)
Implant <- Wrangle(Implant)
Injection <- Wrangle(Injection)
IUD <- Wrangle(IUD)
IUS <- Wrangle(IUS)
Tablet <- Wrangle(Tablet)

#Create Long Acting Reversible Contraception
LARC <- full_join(Implant, Injection)
LARC <- full_join(LARC, IUD)
LARC <- full_join(LARC, IUS)
```

```{r PlottingLARCData, warning=FALSE, message=FALSE}
# Plot per Health Board (Panel A)
#Initialise Data for Health Boards
LARCPlotData <- LARC %>%
  group_by(PaidDateMonth, HealthBoardCode) %>%
  summarise(SumPaidQuantity = sum(PaidQuantity),
            .groups = 'drop')

HealthBoardLabels <-
  c(
    "NHS Ayrshire and Arran",
    "NHS Borders",
    "NHS Dumfries and Galloway",
    "NHS Fife",
    "NHS Forth Valley",
    "NHS Grampian",
    "NHS Greater Glasgow and Clyde",
    "NHS Highland",
    "NHS Lanarkshire",
    "NHS Lothian",
    "NHS Orkney",
    "NHS Shetland",
    "NHS Tayside",
    "NHS Western Isles"
  )

names(HealthBoardLabels) <-
  c(
    "AaA",
    "Bor",
    "DaG",
    "Fif",
    "FoV",
    "Gra",
    "GGC",
    "Hig",
    "Lan",
    "Lot",
    "Ork",
    "She",
    "Tay",
    "WIs"
  )
#Building the plot
LARCPlot <- LARCPlotData %>%
  ggplot(aes(x = PaidDateMonth, y = SumPaidQuantity)) +
  annotate(geom = "rect",
    xmin = c(dmy("23/03/2020"), dmy("26/12/2020"), dmy("26/12/2021")),
    xmax = c(dmy("19/07/2020"), dmy("16/04/2021"), dmy("21/03/2022")),
    ymin = c(-Inf, -Inf, -Inf),
    ymax = c(Inf, Inf, Inf),
  fill = "yellow",
  alpha = 0.25) +
  geom_line(mapping = aes(colour = HealthBoardCode)) +
  geom_point(mapping = aes(colour = HealthBoardCode),
             shape = 1) +
  scale_x_date() +
  facet_wrap(
    ~ HealthBoardCode,
    ncol = 2,
    scales = "free",
    labeller = labeller(HealthBoardCode = HealthBoardLabels)
  ) +
  theme_minimal() +
  scale_color_discrete_qualitative(palette = "Pastel1") +
  labs(title = "Long Acting Reversible Contraception Prescriptions \n Issued by NHS Scotland by Health Board", x = "Month and Year Prescription Dispensed", y = "Sum of Prescriptions Issued") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"), strip.text = element_text(face = "italic"))

#Output = 
#LARCPlot

#Now to build plot for all of Scotland (Panel B)
LARCPlotDataScotland <- LARC %>%
  group_by(PaidDateMonth) %>%
  summarise(SumPaidQuantity = sum(PaidQuantity),
            .groups = 'drop')

LARCPlotScotland <- LARCPlotDataScotland %>%
  ggplot(aes(x = PaidDateMonth, y = SumPaidQuantity)) +
  annotate(geom = "rect",
    xmin = c(dmy("23/03/2020"), dmy("26/12/2020"), dmy("26/12/2021")),
    xmax = c(dmy("19/07/2020"), dmy("16/04/2021"), dmy("21/03/2022")),
    ymin = c(-Inf, -Inf, -Inf),
    ymax = c(Inf, Inf, Inf),
  fill = "yellow",
  alpha = 0.25) +
  geom_line(colour = "#0391BF") +
  geom_point(colour = "#0391BF",
             shape = 1) +
  scale_x_date() +
  theme_minimal() +
  labs(title = "Long Acting Reversible Contraception Prescriptions \n Issued by NHS Scotland", x = "Month and Year Prescription Dispensed", y = "Sum of Prescriptions Issued") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"), strip.text = element_text(face = "italic"))
#Output = 
#LARCPlotScotland

#Now to build plot for all of Scotland but zoomed in on 2020 to 2022 (Panel C)
LARCPlotDataScotlandZoomed <- LARC %>%
  group_by(PaidDateMonth) %>%
  filter(as_date(PaidDateMonth) >= as_date("2020-01-01")) %>% 
  summarise(SumPaidQuantity = sum(PaidQuantity),
            .groups = 'drop')

LARCPlotScotlandZoomed <- LARCPlotDataScotlandZoomed %>%
  ggplot(aes(x = PaidDateMonth, y = SumPaidQuantity)) +
  annotate(geom = "rect",
    xmin = c(dmy("23/03/2020"), dmy("26/12/2020"), dmy("26/12/2021")),
    xmax = c(dmy("19/07/2020"), dmy("16/04/2021"), dmy("21/03/2022")),
    ymin = c(-Inf, -Inf, -Inf),
    ymax = c(Inf, Inf, Inf),
  fill = "yellow",
  alpha = 0.25) +
  geom_line(colour = "#0391BF") +
  geom_point(colour = "#0391BF",
             shape = 1) +
  scale_x_date(date_breaks="2 months", date_labels = "%b '%y", minor_breaks = NULL) +
  theme_minimal() +
  labs(title = "Long Acting Reversible Contraception Prescriptions \n Issued by NHS Scotland from Jan 2020 to Jul 2022", x = "Month and Year Prescription Dispensed", y = "Sum of Prescriptions Issued") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, face = "bold"), axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"), strip.text = element_text(face = "italic"), axis.text.x.bottom = element_text(angle = 45, margin = margin(t = 10)))
#Output = 
#LARCPlotScotlandZoomed

#Now to put A, B & C together to make one plot with `patchwork`
LARCPlotsPatchwork <- wrap_plots(LARCPlot, 
           LARCPlotScotland / LARCPlotScotlandZoomed) +
  plot_layout(widths = c(2,1)) + 
  plot_annotation(caption = 'Disclaimer: Scales are free on these plots. \n Yellow shaded areas indicate national lockdowns in Scotland.', tag_levels = 'A', title = "Change in Prescribing Rates of Long Acting Reversible Contraception in Scotland over Covid-19 Lockdowns", theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 18))) & 
  theme(plot.tag = element_text(face = "bold"), plot.margin = margin(5, 5, 5, 5))
```

\newpage 
\startlandscape

```{r fig.width=11, fig.height=8}
#Output =
LARCPlotsPatchwork
```

\stoplandscape
\newpage 

blah blah